---
title: "STAT 462 Final Project"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# Load necessary Libraries
library(tidyverse)
library(ggplot2)
library(janitor)
library(kableExtra)

# Load Dataset 
data <- read.csv("insurance.csv")
```

# Group Memeber

Eric Wu, Derek Avery, Allison Schaedler, Xinyi Bao

# Background

The Insurance dataset extracted from Kaggle.com describes the medical costs for over 1300 individuals based on their age, sex, BMI, how many children, Smoker or not, and which region in the U.S they lived. The goal of the project is to compare multiple ML method including multiple linear regression in order to find the best model with the lowest MSE value. In addition, there will be a detail diagnostic on the linear regression model for the assumptions, outliers, and other possible issues. 

# Exploratory Data Analysis
```{r}
# Check for Null Value
na <- sum(is.na(data))

# Create dummy variable
dummy_data <- data %>%
  mutate(
    is_male = if_else(sex == "male", 1, 0),
    is_smoker = if_else(smoker == "yes", 1, 0),
    is_southwest = if_else(region == "southwest", 1, 0),
    is_southeast = if_else(region == "southeast", 1, 0),
    is_northwest = if_else(region == "northwest", 1, 0),
    is_northeast = if_else(region == "northeast", 1, 0)
  ) %>%
  select(-sex, -smoker, -region)

# Create standardized dataset for ML that requires scaling
scaled_data <- scale(dummy_data)
```

## Data Statistics
```{r}
# Create a basic statistic table using the quantitative variables of the dataset.
quant <- c("age", "bmi", "children", "charges")

summary_stat <- psych::describe(data[quant], skew = FALSE) %>%
  round(2)

summary_stat %>%
  kable(
    caption = 'Basic statistics of quantitative variables',
    booktabs = TRUE,
    align = c('l', rep('c', 8))
  ) %>%
  kableExtra::kable_styling(
    bootstrap_options = c('striped', 'condensed'),
    font_size = 10,
    latex_options = "hold_position"
  ) 
```

## Data Visualization

# Model Building

## Multiple Linear Regression

## LINE Assumption

The basic assumptions for a linear regression are: Linearity, Independence, Normality, and Equivariance. If any of these assumptions are violated, there are two concerns: incorrect conclusions and interpretations, or a lack of power for discovery.

We analyze these assumptions by using diagnostic plots on our full model.

```{r}
# Full linear model using bmi as response
lm_fullModel<-lm(dummy_data$bmi~., data=dummy_data)
summary(lm_fullModel)

# Residuals against fitted values plot
plot(lm_fullModel$fitted.values, lm_fullModel$residuals)

# QQ Test for normality
qqnorm(lm_fullModel$residuals)
qqline(lm_fullModel$residuals)

# Shapiro - Wilk Test for normality
shapiro.test(lm_fullModel$residuals)
```

Based on the Shapiro-Wilk test, the data is not normally distributed. However, due to the fact our dataset is quite large, we can reasonably ignore the issue. This is because (n - p) >= 30 where (1338 - 10) >= 30.

## KNN Regression

## Random Forest

# Model Comparison

# Discussion

# Conclusion

# Code Appendix 


